---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Travis build status](https://travis-ci.org/flostracke/efor.svg?branch=master)](https://travis-ci.org/flostracke/efor)
[![Coverage status](https://codecov.io/gh/flostracke/efor/branch/master/graph/badge.svg)](https://codecov.io/github/flostracke/efor?branch=master)
[![stability-experimental](https://img.shields.io/badge/stability-experimental-orange.svg)](https://github.com/joethorley/stability-badges#experimental)


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# efor

The goal of efor is to make it easier if you have to forecast or apply preprocessing to multiple timeseries.
Furthermore the the package supports in transforming timeseries data for applying machine learning or deeplearning to it.
(Creation of lagged values, boxcox transformation)

## Installation

You can install the released version of efor from github with:

``` r
#devtools::install_github("flostracke/efor")
```

## Example

First we load some packages for this example.
```{r message=FALSE, warning=FALSE}

library(salesdata) # devtools::install_github("flostracke/salesdata"). Some example data
library(tidyverse)
library(efor)
library(furrr) # for running the forecasting in parallel
library(forecast) #provides forecast mehotds
library(prophet) # provides forecast mehod
library(tidyquant) # for nicer plots

sales_data <- sales_monthly
```

We have some sales data for four articles. We want to create forecasts for all these articles.
The efor package makes this quite easy, because it provides functionality to create forecasts
for multiple articles with one line of code. Right now you can use the forecast package or 
the prophet package.

```{r}

ggplot(sales_data, aes(x = date, y = y)) +
  geom_line() +
  geom_point() +
  facet_wrap(~iterate) +
  ggtitle("The original series") +
  theme_tq() 
```

We split the dataset in a train and test set. All observations from the year 2016
go to the test set. We want to create forecasts for the next 4 months of the testset 
and evaluate the performance.

```{r}
train_data <- sales_data %>% 
  filter(date < "2016-01-01")

test_data <- sales_data %>% 
  filter(date >= "2016-01-01")
```

Now we can apply the the auto.arima function to the dataset and create the forecasts.
```{r}
forecasts_ar <- tf_grouped_forecasts(
  train_data, # used training dataset
  n_pred = 4, # number of predictions
  func = auto.arima, # used forecasting method
  freq = 12 # Frequency. 12 for monthly data, 1 for daily data
)
```

With the same syntax you can create forecasts utilizing the prophet package:
Please note that there we disable the parallel function for prophet, because there 
is some bug right now.

```{r message=FALSE, warning=FALSE}
forecasts_prophet <- tf_grouped_forecasts(
  train_data, # used training dataset
  n_pred = 4, # number of predictions
  func = prophet, # used forecasting method
  freq = 12, # Frequency. 12 for monthly data, 1 for daily data
  parallel = FALSE
)
```

In order to create some plots and evaluate the performance we combine the forecasts
into one dataset.

```{r}
forecasts <- bind_rows(forecasts_ar, forecasts_prophet)
```

The package brings also a function which makes ist quite easy to access the performance (right now thhe mae, rmse and rsquared is calculated)of multiple methods:

```{r}
tf_calc_metrics(forecasts, test_data)
```

Also it is possible to access the performance of each article:

```{r}
tf_calc_metrics(forecasts, test_data, detailed = TRUE)
```

Finally we create a quick graph visualising the results of the forecasts.

```{r}
train_data_plot <- train_data %>% 
  mutate(key = "train")

test_data_plot <- test_data %>% 
  mutate(key = "test")

bind_rows(train_data_plot, test_data_plot) %>% 
  bind_rows(forecasts) %>% 
  ggplot(aes(x = date, y = y, color = key)) +
  geom_point() +
  geom_line() +
  facet_wrap(~iterate) +
  ggtitle("Forecasted values for each article") +
  ylab("Sales amount") +
  theme_tq() +
  scale_fill_viridis(option="magma")
```

And we can plot the forecasts against the actual forecasts:
```{r}
tf_plot_preds_actuals(forecasts, test_data)
```


