---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Travis build status](https://travis-ci.org/flostracke/efor.svg?branch=master)](https://travis-ci.org/flostracke/efor)
[![Coverage status](https://codecov.io/gh/flostracke/efor/branch/master/graph/badge.svg)](https://codecov.io/github/flostracke/efor?branch=master)
[![stability-experimental](https://img.shields.io/badge/stability-experimental-orange.svg)](https://github.com/joethorley/stability-badges#experimental)


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# efor

The goal of EasyFORcasting or short efor is to make it easier if you have to forecast or apply preprocessing to multiple timeseries.
Furthermore the the package supports in transforming timeseries data for applying machine learning or deeplearning to it.
(Creation of lagged values)

## Installation

You can install the released version of efor from github with:

``` r
#devtools::install_github("flostracke/efor")
```

## Example

First we load some packages for this example.
```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(tsibble) # for creating nicer representation of monthly data
library(efor)
library(furrr) # for running the forecasting in parallel
library(forecast) #provides forecast mehotds
library(prophet) # provides forecast mehod


sales_data <- sales_monthly %>% 
  mutate(date = yearmonth(date))

sales_data
```

We have some sales data for four articles. We want to create forecasts for all 
these articles.
The efor package makes this quite easy, because it provides functionality to 
create forecasts for multiple articles with just one function code. Right now 
you can apply the functions from the forecast or prophet package.

The idea is that all the data has to be organised in one dataframe with the 
following columns:

- date: A date column
- iterate: the grouping variable. In this example it is the articlenumber
- y: the value you want to forecast

The dataframe sales_data is already meeting these requirements.

Before we start forecasts let's quickly create a plot of the data:

```{r}

ggplot(sales_data, aes(x = date, y = y)) +
  geom_line() +
  geom_point() +
  facet_wrap(~iterate) +
  ggtitle("The original series") +
  theme_minimal() 
```

We split the dataset in a train and test set. All observations from the year 2016
go into the test set. We want to create forecasts for the next 4 months of the 
testset and evaluate the performance of different methods.

```{r}
train_data <- sales_data %>% 
  filter(date < "2016-01-01")

test_data <- sales_data %>% 
  filter(date >= "2016-01-01")
```

Now we can apply the the auto.arima function to the dataset and create the 
forecasts. All the methods from the forecast package can be run in parallel.

```{r}
forecasts_ar <- tf_grouped_forecasts(
  train_data,        # used training dataset
  n_pred = 6,        # number of predictions
  func = auto.arima, # used forecasting method
  parallel = TRUE    # for runiing in parallel
)

forecasts_ar
```

With the same syntax you can create forecasts utilizing the prophet package:
Please note that there we disable the parallel function for prophet, because there 
is some bug right now.

```{r message=FALSE, warning=FALSE}
forecasts_prophet <- tf_grouped_forecasts(
  train_data,      # used training dataset
  n_pred = 6,      # number of predictions
  func = prophet,  # used forecasting method
  parallel = FALSE #disabling parallel for prohet
)

forecasts_prophet
```

In order to create some plots and evaluate the performance we combine the forecasts
into one dataset.

```{r}
forecasts <- bind_rows(forecasts_ar, forecasts_prophet) %>% 
  mutate(date = yearmonth(date)) #reformat the date because of a bug in bind_rows
```

The package brings also a function which makes ist quite easy to access the performance (right now thhe mae, rmse and rsquared is calculated)of multiple methods:

```{r}
tf_calc_metrics(forecasts, test_data)
```

Also it is possible to access the performance of each article:

```{r}
tf_calc_metrics(forecasts, test_data, detailed = TRUE)
```

Finally we create a quick graph visualising the results of the forecasts.

```{r}
train_data_plot <- train_data %>% 
  mutate(key = "train")

test_data_plot <- test_data %>% 
  mutate(key = "test")

bind_rows(train_data_plot, test_data_plot) %>% 
  bind_rows(forecasts) %>% 
  filter(key %in% c("auto.arima", "train", "test")) %>% 
  mutate(date = yearmonth(date)) %>% 
  ggplot(aes(x = date, y = y, color = key)) +
  geom_point() +
  geom_line() +
  facet_wrap(~iterate) +
  ggtitle("Forecasted values for each article") +
  ylab("Sales amount") +
  theme_minimal()
```


 
 



