---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Travis build status](https://travis-ci.org/flostracke/efor.svg?branch=master)](https://travis-ci.org/flostracke/efor)
 [![Coverage status](https://codecov.io/gh/flostracke/tidypreprocess/branch/master/graph/badge.svg)](https://codecov.io/github/flostracke/tidypreprocess?branch=master)

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# tidypreprocess

The goal of tidypreprocess is to make it easier if you have to apply preprocessing to multiple timesereis

## Installation

You can install the released version of tidypreprocess from github with:

``` r
devtools::install_github("flostracke/tidypreprocess")
```

## Example

First we load some packages for this example.
```{r}

library(salesdata) # devtools::install_github("flostracke/salesdata")
library(tidyverse)
library(tidypreprocess)

sales_data <- sales_monthly
```

We have some sales data for four articles. Prior to forecasting to these series
we want to apply the boxcox transformation. This makes the series normal 
distributed.

```{r}

ggplot(sales_data, aes(x = date, y = y)) +
  geom_line() +
  geom_point() +
  facet_wrap(~iterate) +
  ggtitle("The original series")
```

We retrieve the lambdas for the dataset. With these lambdas we can revert
the boxcox transformation.

```{r}
lambdas <- tf_get_lambdas(sales_data)

lambdas
```

Apply the boxcox transformation

```{r}

sales_data_trans <- tf_boxcox(sales_data)
```

A quick plot of the transformed data:
```{r}
ggplot(sales_data_trans, aes(x = date, y = y)) +
  geom_line() +
  geom_point() +
  facet_wrap(~iterate) +
  ggtitle("The boxcox transformed series")
```

Now we remove the transformation

```{r}

sales_data_retransformed <- tf_remove_boxcox(sales_data_trans, lambdas) %>%

  # we round the numbers in order to compare for resulting differences due to the
  # backtransformation
  mutate(y = round(y,0))
```

The transformation and the backtransformation is working

```{r}
sales_data %>%
  left_join(
    sales_data_retransformed,
    by = c("date", "iterate"),
    suffix = c("_orig", "_trans")
  ) %>%
  mutate(diff = y_orig - y_trans) %>%
  filter(diff != 0)
```






